Tulai

### 2023-11-05 paper/paper.qmd
add map of sites, update stats in text, update table of post-hoc tests, update PCA plot

```{r}
# Read the data
tl <- read_csv(here::here("data/raw-data/Tulai_Lithic_Assamblage.csv"))

# this file is prepared from the google sheet here
# https://docs.google.com/spreadsheets/d/1Q0QZESk412ZQLE24yPs6Rg-7Y9OMX63DHoXeIPNAIFM/edit#gid=0
# and simplified using code in the scraps.R file
excavation_data <- 
  read_csv(here::here("data//raw-data/excavation_data.csv"))
```

@@

```{r}
library(ggrepel)

@@

# get excavation area so we can use it for grouping
excavation_area <- 
  str_extract(sum_depth_sheet_subset$label2, 
              "[^_]+")
pca_data.pca <- prcomp(pca_data, scale. = TRUE)

@@

pca_data.pca <- prcomp(pca_data, scale. = TRUE)

biplot_modified <- 
ggbiplot(
  pca_data.pca,
  obs.scale = 1,
  var.scale = 1,
  groups = excavation_area,
  ellipse = FALSE,
) +
  stat_chull(aes(color = excavation_area,
                 fill = excavation_area),
             alpha = 0.1,
             geom = "polygon") +
  scale_colour_brewer(palette = "Set1",
                      name = '',
                      guide = 'none') +
  scale_fill_brewer(palette = "Set1",
                    name = '')  +
  theme_minimal()

@@

>>>>

```{r}
# Read the data
tl <- read_csv(here::here("data/raw-data/Tulai_Lithic_Assamblage.csv"))
site_locations <- read_csv(here::here("data/raw-data/Sites-mentioned-in-the-text.csv"))

# this file is prepared from the google sheet here
# https://docs.google.com/spreadsheets/d/1Q0QZESk412ZQLE24yPs6Rg-7Y9OMX63DHoXeIPNAIFM/edit#gid=0
# and simplified using code in the scraps.R file
excavation_data <- 
  read_csv(here::here("data/raw-data/excavation_data.csv"))
```

Draw the map and save it

```{r}
#| eval: false

# get map tiles from the internet, we only need to do this once and save them

library(ggmap) # devtools::install_github("stadiamaps/ggmap")

# using http://bboxfinder.com/ to get this
bbox = c(
  42, # left
  30, # lower
  51, # right
  38) # upper

# this takes a few minutes and needs an internet connection
map <-
  get_stadiamap(bbox = bbox,
                zoom = 7)

saveRDS(map, file = here::here("data/raw-data/map_tiles.rds"))

# map with site names next to the points
map_of_sites <-
  ggmap(map)  +
  geom_point(data = site_locations,
             aes(y = `Longitude DD` ,
                 x = `Latitude DD`),
             colour = "red",
             size = 2) +
  # label points with site name
  geom_text_repel(data = site_locations,
             aes(y = `Longitude DD` ,
                 x = `Latitude DD`,
                      label = Site),
                  size = 2,
                  bg.color = "white",
                  bg.r = 0.15) +
  theme_minimal(base_size = 6) +
  labs(x = "Longitute",
       y = "Latitude") +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())

# inset world map, from
# https://r-spatial.org/r/2018/10/25/ggplot2-sf.html
library(sf)
library("rnaturalearth")
library("rnaturalearthdata")

world <- ne_countries(scale = "medium", returnclass = "sf")

worldplot <- 
ggplot(data = world) +
     geom_sf(color = "black", 
             fill = "white",
             linewidth = 0.05) +
    annotate("rect",
           xmin = bbox[3],
           xmax = bbox[1],
           ymin = bbox[2],
           ymax = bbox[4],
           fill = NA,
           colour = "red",
           size = 1.5
  ) +
  theme_void() +
  theme(panel.background = element_rect(fill = "white"))

# combine site map and inset map
library(cowplot)
ggdraw() +
  draw_plot(map_of_sites) +
  draw_plot(plot = worldplot,
            x = -0.02, # x location of inset placement
            y = -0.06, # y location of inset placement
            width = .6, # Inset width
            height = .4, # Inset height
            scale = 0.75 # Inset scale
  ) +
  theme(panel.background = element_rect(fill='white', colour="white"),
        plot.background = element_rect(fill='white', colour="white"))

ggsave(here::here("figures/site-map.png"),
       width = 4.45,
       height = 4.75,
       units = "in")

Display the map

```{r}
#| label: fig-map
#| out-width: "400px"
knitr::include_graphics(here::here("figures/site-map.png"))
```

@@

```{r}
#| label: fig-pca-biplot

library(ggrepel)

@@

# get excavation area so we can use it for grouping
excavation_area <- 
  str_extract(sum_depth_sheet_subset$label2, 
              "[^_]+")

@@

names(pca_data) <- str_replace_all(names(pca_data), "_", " ")

pca_data.pca <- prcomp(pca_data, scale = TRUE)

biplot_modified <- 
ggbiplot(
  pca_data.pca,
  obs.scale = 1,
  var.scale = 1,
  groups = excavation_area,
  ellipse = FALSE,
) +
  stat_chull(aes(color = excavation_area,
                 fill = excavation_area),
             alpha = 0.1,
             geom = "polygon") +
  scale_colour_brewer(palette = "Set1",
                      name = '',
                      guide = 'none') +
  scale_fill_brewer(palette = "Set1",
                    name = '')  +
  scale_x_continuous(limit = c(-6, 3.5)) +
  scale_y_continuous(limit = c(-3, 3)) +
  theme_minimal() +
  theme(legend.position = c(0.1, 0.9),
        legend.background = element_rect(fill = "white", linewidth = 0))

@@

set.seed(0411)
